Formalizing Stack Safety as a Security Property

### 2023-02-03 submission5/core.tex

Computer programming relies on functions (or subroutines, procedures, methods, etc.)
as a foundational abstraction. Functions are units of computation that
can call one another to realize larger computations in a modular fashion.
%
At a low level, a function activation manages its own data, such as local variables, as
well as information about its \emph{caller} function, to which it must return.
The \emph{(call) stack} is the fundamental data structure charged with realizing
this abstraction in memory. \an{registers?} From a security perspective, attacks on the stack are really attacks
on the fundamental functional abstraction itself. %, and are therefore every bit as
%pervasive.

Indeed, the call stack is a ancient~\cite{one06phrack} and perennial~\cite{mitre-cwe,DBLP:conf/raid/VeendCB12,
  DBLP:conf/sp/SzekeresPWS13,
  DBLP:conf/sp/HuSACSL16,msrc-bluehat,chromium-security}
  target for low-level attacks. \apt{characterize the variety of attacks?}
%
%The most recent release of the authoritative CWE Top 25
%Most Dangerous Software Weaknesses~\cite{
%list\footnote{\url{https://cwe.mitre.org/top25/archive/2022/2022_cwe_top25.html}}
%shows that general classes of vulnerabilities closely related to the stack
%% \rb{(though not exclusively)}
%consistently rank at and near the top of the list, e.g., \#1 Out-of-bounds Write, \#5 Out-of-bounds Read, and \#7 Use After Free.
%
%Moreover, many other types of faults involve some kind of control or data flow
%using data stored in the stack 
The variety in attacks on the stack is mirrored by the
software and hardware protections that aim to prevent them,
%
including stack canaries~\cite{Cowan+98},
bounds checking~\cite{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
split stacks~\cite{Kuznetsov+14},
shadow stacks~\cite{Dang+15,Shanbhogue+19},
capabilities~\cite{Woodruff+14,Chisnall+15,SkorstengaardLocal,SkorstengaardSTKJFP,Georges22:TempsDesCerises},
and hardware tagging~\cite{DBLP:conf/sp/RoesslerD18,Gollapudi+23}.

\apt{Add brief case for having formal guarantees.}
Of these mechanisms, many are fundamentally ill-suited for offering formal guarantees,
as they aim to impede attackers but cannot provide universal protection. Shadow stacks,
for instance, aim to ``restrict the flexibility available in creating gadget chains''
\cite{Shanbhogue+19}, not to categorically
rule out attacks. Other mechanisms are stronger, but formal
guarantees are rarely discussed.
Indeed, the Cerise work of Skorstengaard et al. \cite{SkorstengaardSTKJFP}\apt{that's not Cerise} and
Georges et al. \cite{Georges22:TempsDesCerises} is so far the sole line of work
to offer (various)\bcp{Maybe we can omit that word here and spell it
  out in the related work.} stack safety formalizations.

\apt{Transition}
Roessler and DeHon \cite{DBLP:conf/sp/RoesslerD18} present an array of tag-based
micro-policies~\cite{pump_oakland2015} for stack safety, at least some of which should be amenable to making
formal guarantees. Among these is their ``Lazy Tagging and Clearing'' (\(LTC\)) micro-policy,
which is is peculiar in that function activations may write into one another's stack frames,
as long as the data cannot be accessed afterward. Roessler and DeHon themselves say that
\(LTC\) does not enforce memory safety, but rather a form of data-flow integrity~\cite{castro+06}.
And \(LTC\) definitely would not enforce a tag-based adaptation of Georges et al.'s property.

Neverthess, we believe that there is a useful sense in which \(LTC\) does (attempt to) enforce stack safety. 
It does not always protect individual addresses on the stack, but it tries to 
prevent one function activation from interfering with another's behavior---protecting 
the functional abstraction, which we argue is the real essence of stack safety.
%We argue that a formal characterization of \(LTC\)'s protections
%should fall under the umbrella of stack safety.

@@

The structure of our properties enables them to be easily extended to systems with
sophisticated features such as callee-save registers, argument-passing on the stack,
and tailcall-elimination. We likewise extend \(LTC\) to protect these features, and
test the extended micro-policy against our property. We find that, apart from the extensions, \(LTC\)
is flawed in a way that undermines integrity and confidentiality, and we correct
this flaw.

@@

Otherwise, we do not assume that the code adheres to any particular
calling convention or implements a source language construct.
In particular, while we are agnostic as to the source
language, we certainly aim to support C, and so any source function might contain undefined
behavior resulting in its compilation to arbitrary machine code. A given enforcement
mechanism may place additional constraints, particularly on the behavior of
call and return sequences. For instance, extant implementations tend to assume
implicitly that callee-saved registers are treated appropriately by whichever compiler
generated their code. Our properties explicitly require them to be treated as belonging
to the caller, which could be enforced by a micro-policy or by a well-behaved compiler.
%
\rb{This last bit about callee-saved registers belonging to the caller is subtle, I think it risks the reader missing the point.}

@@

This is a strong threat model, but it does omit some important aspects of
security enforcement: in particular, hardware and timing attacks are out of scope, and
we do not address concurrency.
%
\rb{Anything about concurrency in future work, though?}

>>>>

Computer programming relies on functions (or subroutines, procedures, methods, etc.)
as a foundational abstraction. Functions are units of computation that
can call one another to realize larger computations in a modular fashion.
%
At a low level, a function activation manages its own data, such as local variables, as
well as information about its \emph{caller} function, to which it must return.
The \emph{(call) stack} is the fundamental data structure charged with realizing
this abstraction, aided by an ABI that defines how registers are shared between activations.
From a security perspective, attacks on the stack are really attacks
on the fundamental functional abstraction itself, and are therefore every bit as
pervasive.

Indeed, the call stack is a ancient~\cite{one06phrack} and
perennial~\cite{mitre-cwe,DBLP:conf/raid/VeendCB12,
  DBLP:conf/sp/SzekeresPWS13,
  DBLP:conf/sp/HuSACSL16,msrc-bluehat,chromium-security}
target for low-level attacks, frequently involving control-flow hijacking via corrupting
the return address, and memory corruption more generally.

%
%The most recent release of the authoritative CWE Top 25
%Most Dangerous Software Weaknesses~\cite{
%list\footnote{\url{https://cwe.mitre.org/top25/archive/2022/2022_cwe_top25.html}}
%shows that general classes of vulnerabilities closely related to the stack
%% \rb{(though not exclusively)}
%consistently rank at and near the top of the list, e.g., \#1 Out-of-bounds Write, \#5 Out-of-bounds Read, and \#7 Use After Free.
%
%Moreover, many other types of faults involve some kind of control or data flow
%using data stored in the stack 
The variety in attacks on the stack is mirrored by the
software and hardware protections that aim to prevent them,
%
including stack canaries~\cite{Cowan+98},
bounds checking~\cite{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
split stacks~\cite{Kuznetsov+14},
shadow stacks~\cite{Dang+15,Shanbhogue+19},
capabilities~\cite{Woodruff+14,Chisnall+15,SkorstengaardLocal,SkorstengaardSTKJFP,Georges22:TempsDesCerises},
and hardware tagging~\cite{DBLP:conf/sp/RoesslerD18,Gollapudi+23}.

\apt{Add brief case for having formal guarantees.}
Of these mechanisms, many are fundamentally ill-suited for offering formal guarantees,
as they aim to impede attackers but cannot provide universal protection. Shadow stacks,
for instance, aim to ``restrict the flexibility available in creating gadget chains''
\cite{Shanbhogue+19}, not to categorically
rule out attacks. Other mechanisms such as Softbound\cite{NagarakatteZMZ09} and Code-pointer
integrity\cite{Kuznetsov+14} are stronger, but most do not give formal guarantees.
Indeed, the secure calling conventions of Skorstengaard et al. \cite{SkorstengaardSTKJFP} and
Georges et al. \cite{Georges22:TempsDesCerises} is so far the sole line of work
to offer stack safety formalizations.

Among the mechanisms that should be amenable to making strong formal guarantees,
Roessler and DeHon \cite{DBLP:conf/sp/RoesslerD18} present an array of tag-based
micro-policies~\cite{pump_oakland2015} for stack safety.
Their most realistic micro-policy, ``Lazy Tagging and Clearing'' (\(LTC\)),
makes an interesting trade-off for performance: it allows function activations to write
into one another's stack frames, but ensures that the owner of the corrupted memory cannot access
it afterward. Because the memory does get corrupted, \(LTC\) would not fulfill an
adaptation Georges et al.'s property to the tagged setting.

Neverthess, we believe that there is a useful sense in which \(LTC\) does (attempt to) enforce stack safety. 
It does not always protect individual addresses on the stack, but it tries to 
prevent one function activation from interfering with another's behavior---protecting 
the functional abstraction, which we argue is the real essence of stack safety.

@@

The structure of our properties enables them to be easily expanded to characterize systems with
sophisticated features such as callee-save registers, argument-passing on the stack,
and tailcall-elimination. We likewise extend \(LTC\) to protect these features, and
test the extended micro-policy against our property. We find that, apart from the extensions, \(LTC\)
is flawed in a way that undermines integrity and confidentiality, and we correct
this flaw.

@@

Otherwise, we do not assume that the code adheres to any particular
calling convention or implements a source language construct.
In particular, while we are agnostic as to the source
language, we certainly aim to support C, and so any source function might contain undefined
behavior resulting in its compilation to arbitrary machine code. A given enforcement
mechanism or target architecture might place additional constraints, particularly on the behavior of
call and return sequences. For instance, extant implementations tend to assume
implicitly that callee-saved registers have their values maintained by whichever compiler
generated their code. Our properties explicitly state this as a requirement,
which could be enforced by a micro-policy, a well-behaved compiler, or other enforcement technique.

@@

This is a strong threat model, but it does omit some important aspects of
security enforcement: in particular, hardware and timing attacks are out of scope, and
we do not address concurrency---that is a high-priority future work.
%
